---
title: "Module 3 Data 608"
author: "Uliana Plotnikova"
date: "March 2, 2021"
output: 
    prettydoc::html_pretty:
    theme: lumen
    highlight: github
    math: katex
    code_folding: hide
runtime: shiny
---


#### I have provided you with data about mortality from all 50 states and the District of Columbia.

#### Please access it athttps://github.com/charleyferrari/CUNY_DATA608/tree/master/module3/data 





```{r, echo=FALSE}

suppressWarnings ({library(shiny)})
suppressWarnings ({library(rsconnect)})
suppressWarnings ({library(ggplot2)})
suppressWarnings ({library(dplyr)})
suppressWarnings ({library(plotly)})
suppressWarnings ({library(kableExtra)})
suppressWarnings({library(tidyverse)})
suppressWarnings({library(gganimate)})
suppressWarnings({library(shinyWidgets)})

```




```{r}
# load the dataset in R and review
df <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv")
kable(head(df,5))
```
```{r}
summary(df)
```



```{r}
# Mortality Causes
causes <- unique(df$ICD.Chapter)
causes
```




####  Ave, Min, Max Crude rate for each state

```{r}
statedeath<- df %>%
    group_by(State) %>%
    summarise(Avg_Rate = mean(Crude.Rate), 
              Min_Rate = min(Crude.Rate),
              Max_Rate = max(Crude.Rate)) %>%
              
  arrange(State)
kable(head(statedeath,10))
```


#### Question 1: 
#### As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortalityfor each cause of death.



#### Let's explore Neoplasms



```{r}
icd<-filter(df, df$ICD.Chapter== "Neoplasms")
kable(head(icd))
```



```{r}
p <- ggplot(
  icd, 
  aes(x = Crude.Rate, y=Deaths,  colour = State)
  ) +
  geom_point(show.legend = FALSE, alpha = 0.9) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  labs(x = "Crude.Rate", y = "Deaths")
```



```{r}
p + transition_time(Year)+
  labs(title = "Year: {frame_time}")+
    shadow_mark(alpha = 0.3, size = 0.9)
```





```{r}
p <- ggplot(
  icd,
  aes(Year, Crude.Rate, group = ICD.Chapter, color = factor(ICD.Chapter))
  ) + geom_line() +
  scale_color_viridis_d() +
  labs(x = "Year", y = "Crude.Rate") +
  theme(legend.position = "top")
```


```{r}
p + geom_point(aes(group = seq_along(Year))) +
  transition_reveal(Year)
```


#### Now let's take a look at 2010

```{r}

year<-filter(icd, icd$Year == "2010")%>% arrange(Crude.Rate)
kable(tail((year),10))
```





```{r}
ggplot(aes(x =  reorder(State,-Crude.Rate), y = Crude.Rate), data = year) + 
  geom_bar(stat = 'identity',width=0.7, color="#1F3552", fill="darkgreen", 
           position=position_dodge()) + coord_flip()  +
  ylab(' Crude Rate') +
  xlab(' State') +
  ggtitle('Crude Rate by State') +
  geom_text(aes(label=Crude.Rate),
            size=3,
            hjust=-1,
            color="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 65, hjust = 1))
```










```{r}
ui <- fluidPage( 
     setBackgroundColor(color = c("#F7FBFF", "#2171B5"),
    gradient = "linear",
    direction = "bottom"),
  titlePanel('State Moratlity Rate by Cause'),
  sidebarPanel(selectInput('Cause', 'Cause', unique(df$ICD.Chapter),
                selected='Neoplasms')),
  mainPanel(htmlOutput(outputId = 'selection'),
            plotOutput('plot1', height="auto"),
            h6("The Crude Mortality Rate across all States"))
           )

```



```{r}
server <- shinyServer(function(input, output, session) {
  selectedData <- reactive({
    df %>% filter(ICD.Chapter == input$Cause & Year == 2010 )
  })

output$selection <- renderText({
    paste('<b> Crude rate for: </b>', input$Cause)
  })
  
  output$plot1 <- renderPlot({
    
    ggplot(selectedData(), aes(x=reorder(State, -Crude.Rate), y=Crude.Rate)) +
      geom_col(fill = "darkgreen") +
      coord_flip() +
      geom_text(aes(label=Crude.Rate),
                size=3,
                hjust=-0.2,
                color="darkgreen") +
      xlab("State") +
      ylab("Crude Rate") +
      theme(panel.background = element_blank())
  }, height = function() {
    session$clientData$output_plot1_width}
  )
})

shinyApp(ui = ui, server = server)
```



#### Question 2
#### Often you are asked whether particular States are improving their mortality rates (per cause)faster than, or slower than, the national average. Create a visualization that lets your clientssee this for themselves for one cause of death at the time. Keep in mind that the nationalaverage should be weighted by the national population.


```{r}

each_state <- df %>% select(ICD.Chapter, Year, Crude.Rate, State)
```



```{r}

nation<- df %>% select(ICD.Chapter, Year, State, Deaths, Population, Crude.Rate) %>%  mutate(Crude.Rate = round((sum(Deaths) / sum(Population)) * 10^5, 1))
```




```{r}
nation_ave <- nation %>%
  select(ICD.Chapter, Year, Crude.Rate, State) %>%
  mutate(State = "National_Ave") %>%
  group_by(ICD.Chapter, Year)
head (nation_ave,10)
```







```{r}
total <- union_all(each_state, nation_ave)
```






```{r}
ui <- fluidPage(
    sidebarPanel(
    selectInput('State', 'State', unique(total$State), selected='NY'),
    selectInput('ICD.Chapter', 'Cause', unique(total$ICD.Chapter), selected='Neoplasms')
  ),
   setBackgroundColor(color = c("#F7FBFF", "#2171B5"),
    gradient = "linear",
    direction = "bottom"),
   titlePanel("State Mortality Rate VS National Average"),
  mainPanel(width=10,
    plotlyOutput('plot1')
  )
)

server <- function(input, output, session) {
  
  nationalData <- reactive({
    nation_ave %>%
      filter(ICD.Chapter == input$ICD.Chapter)
  })
  
  statedata <- reactive({
    df <- total %>%
      filter(State == input$State, ICD.Chapter == input$ICD.Chapter)
  })
  
  combined  <- reactive({
    merge(x = nationalData(), y = statedata(), all = TRUE)
  })
  
  output$plot1 <- renderPlotly({
    
    df <- total %>%
      filter(State == input$State, ICD.Chapter == input$ICD.Chapter)
    
    plot_ly(combined(), x = ~Year, y = ~Crude.Rate, color = ~State, type='scatter',
            mode = 'markers+lines', line = list(color = '#17BECF'))
  })
  
}

shinyApp(ui,server)
```

